Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2018-10-02 paper/paper.tex
very minor tweaks Results section 5

To more systematically assess viral features associated with infection outcome, we divided the 150 cells in \FIG{genotypes} into those that expressed unmutated copies of all eight genes (disregarding synonymous mutations) and those that did not (due to gene absence, amino-acid substitution, insertion, or deletion).
\FIG{mutations}A shows that the 49 cells infected by full wildtype viruses had a significantly tighter distribution of the amount of viral mRNA per cell than the other 101 cells as quantified by the Gini index~\citep{gini1921measurement}.
Therefore, viral genetic variation is a major contributor to the heterogeneity in viral transcriptional burden among cells.

Viral genetic variation also associated with IFN induction, since cells infected by full wildtype virions expressed less IFN than cells infected by incomplete or mutated virions (\FIG{mutations}B).
Specific viral defects were especially associated with IFN induction (\FIG{mutations}C).
For instance, absence of NS and amino-acid mutations in PB1 were significantly enriched in IFN+ cells, and amino-acid mutations in NS and deletions in HA were more weakly enriched.
Only the absence of NS remained significant at a false discovery rate of 10\% (\FIG{mutations}C)---although a Bayesian-minded virologist might contend that controlling for all the hypotheses in \FIG{mutations}C is too stringent, since some are much more plausible \textit{a priori} given existing biological knowledge~\citep[e.g.,][]{killip2017single, velthuis2018mini, wu2014high}.
In any case, \FIG{genotypes} and \FIG{mutations}C motivate compelling hypotheses about specific viral mutations that contribute to IFN induction, and below we pursue these hypotheses using experimental follow-up.

One other interesting trend emerges from the single-cell data.
There is no difference in the amount of viral mRNA between IFN+ and IFN- cells that express NS---but among cells that lack NS, the IFN+ ones have significantly more viral mRNA (\FIG{mutations}D).
The statistical analysis in \FIG{mutations}D is insufficient to determine if the higher viral transcriptional burden in IFN+ NS-deficient infections is a cause or a consequence of the IFN induction, but the experimental follow-up below sheds more light on this question.

>>>>

To systematically assess viral features associated with infection outcome, we divided the 150 cells in \FIG{genotypes} into those that expressed unmutated copies of all eight genes (disregarding synonymous mutations) and those that did not (due to gene absence, amino-acid substitution, insertion, or deletion).
\FIG{mutations}A shows that the 49 cells infected by full wildtype viruses had a significantly tighter distribution of the amount of viral mRNA per cell than the other 101 cells as quantified by the Gini index~\citep{gini1921measurement}.
Therefore, viral defects are a major contributor to the heterogeneity in viral transcriptional burden among cells.

Viral features also associated with IFN induction, with cells infected by full wildtype virions turning IFN+ less often than cells infected by incomplete or mutated virions (\FIG{mutations}B).
Specific viral defects were especially associated with IFN induction (\FIG{mutations}C).
Absence of NS and amino-acid mutations in PB1 were significantly enriched in IFN+ cells, and amino-acid mutations in NS and deletions in HA were weakly enriched.
Only the absence of NS remained significant at a false discovery rate of 10\% (\FIG{mutations}C)---although a Bayesian-minded virologist might contend that controlling for all the hypotheses in \FIG{mutations}C is too stringent, since some are more plausible \textit{a priori} given existing biological knowledge~\citep[e.g.,][]{killip2017single, velthuis2018mini, wu2014high}.
In any case, \FIG{genotypes} and \FIG{mutations}C motivate compelling hypotheses about specific viral defects that contribute to IFN induction, and below we pursue these hypotheses using experimental follow-up.

One other interesting trend emerges from the single-cell data.
There is no difference in the amount of viral mRNA between IFN+ and IFN- cells that express NS---but among cells that lack NS, the IFN+ ones have significantly more viral mRNA (\FIG{mutations}D).
The statistical analysis in \FIG{mutations}D is insufficient to determine if the higher viral transcriptional burden in IFN+ NS-deficient infections is a cause or a consequence of the IFN induction, but the experiments below shed more light on this question.
